The Array of Long Baseline Antennas for Taking Radio Observations from the Sub-Antarctic

### 2020-08-27 albatros_paper.tex
HCC minor edits

Using the eight coastal huts, the \prizm\ site, and the hydro shack as
the nominal \albatros\ installation locations, the computed
synthesized beam is as shown in \autoref{Fig:marion_beam}.  The beam
width at 5~MHz is roughly \SI{7}{\arcminute}, which represents over an
order of magnitude improvement in resolution over other existing
measurements.  One of the challenges in constructing an interferometer
array on Marion Island is that the rugged terrain precludes the
possibility of directly cabling and correlating antennas across large
distances.  The final \albatros\ antenna stations will therefore
operate {\it autonomously}, recording baseband data over extended
periods of time for subsequent offline correlation.  We have conducted
two engineering runs: 1) a two-element, directly correlated pathfinder
to qualitatively understand the sky signal, and 2) a single station to
test the readout and power handling technology that are required for
autonomous operation.

@@

\section{Single autonomous station pathfinder}

@@

The full \albatros\ configuration will comprise an array of 10
autonomous antenna stations, each recording baseband over tunable
frequency windows (with 10--20~MHz total bandwidth) within the full
0--125~MHz operating range.  The baseband data will be collected
periodically and subsequently correlated offline.  The
\albatros\ stations, located at the eight coastal hut sites plus the
hydro shack and \prizm\ site, will be separated by baselines of
$\sim20$~km as shown in \autoref{Fig:marion_map_beam}. One fully
autonomous \albatros\ station, shown in \autoref{Fig:autonomous}, was
deployed in April 2019 at the hydro shack location (\ang{46;52.205;}S,
\ang{37;50.612;}E) on Marion Island as a first step in testing the
technology needed to establish the full array.  The system block
diagram is shown in \autoref{Fig:albatros1_schem}.  The antenna and
front-end active balun in the single autonomous station are identical
to those used in the two-element pathfinder (\S\ref{s:antenna} and
\S\ref{s:fee}), and the back-end electronics and power system are
described in detail below.

@@

Although the first-deployed two-element pathfinder is located at a
reasonable walking distance from the main base and uses a generator to
periodically recharge batteries, the bulk of the antenna stations that
will comprise the full array will be located at points farther removed
from the main base and will therefore require fully autonomous energy
sources.  For the autonomous station pathfinder, we developed a solar
charging system to power the station, and we are investigating small
wind turbines for future \albatros\ stations.

@@

We have been unable to retrieve data from the single autonomous
station pathfinder on Marion because of COVID-19 restrictions and
limited bandwidth to the island, although we have confirmed that the
solar power system has successfully met the power demands of the
antenna station.  To demonstrate baseband writing capability in the
absence of Marion data, we instead present a short observation that
was taken by a separate, single-element \albatros\ pathfinder station
that operated from the McGill Arctic Research Station (MARS;
\ang{79;26;}N, \ang{90;46;}W) on Axel Heiberg Island, Nunavut, during
July~2019.  The RF signal chain of the MARS installation had the same
architecture as that shown in \autoref{Fig:albatros1_schem}, but the
power was supplied by manually charged batteries, rather than a solar
power system.  The MARS RFI environment is somewhat noisier than that
of Marion Island, and because the data were taken during the height of
Arctic summer, the ionospheric plasma cutoff frequency is higher than
the typical cutoff reflected in the Marion data.  Despite the
differences in the power configuration and RFI environment, the MARS
installation is suitable for illustrating baseband data writing.

>>>>

Using the eight coastal huts, the \prizm\ site, and the hydro shack as
the nominal \albatros\ installation locations, the computed
synthesized beam is shown in \autoref{Fig:marion_beam}.  The beam
width at 5~MHz is roughly \SI{7}{\arcminute}, which represents over an
order of magnitude improvement in resolution over other existing
measurements.  One of the challenges in constructing an interferometer
array on Marion Island is that the rugged terrain precludes the
possibility of directly cabling and correlating antennas across large
distances.  The final \albatros\ antenna stations will therefore
operate {\it autonomously}, recording baseband data over extended
periods of time for subsequent offline correlation.  We have conducted
two engineering runs: 1) a two-element, directly correlated pathfinder
to qualitatively understand the sky signal, and 2) a single station to
test the readout and power handling technology that are required for
autonomous operation.

@@

\section{Single autonomous station pathfinder}\label{s:autonomous}

@@

The full \albatros\ configuration will comprise an array of 10
autonomous antenna stations, each recording baseband over tunable
frequency windows (with 10--20~MHz total bandwidth) within the full
0--125~MHz operating range.  Baseband data will be periodically
physically retrieved from each station and subsequently correlated
offline.  The \albatros\ stations, located at the eight coastal hut
sites plus the hydro shack and \prizm\ site, will be separated by
baselines of $\sim20$~km as shown in
\autoref{Fig:marion_map_beam}. One fully autonomous
\albatros\ station, shown in \autoref{Fig:autonomous}, was deployed in
April 2019 at the hydro shack location (\ang{46;52.205;}S,
\ang{37;50.612;}E) on Marion Island as a first step in testing the
technology needed to establish the full array.  The system block
diagram is shown in \autoref{Fig:albatros1_schem}.  The antenna and
front-end active balun in the single autonomous station are identical
to those used in the two-element pathfinder (\S\ref{s:antenna} and
\S\ref{s:fee}), and the back-end electronics and power system are
described in detail below.

@@

Although the first-deployed two-element pathfinder is located at a
reasonable walking distance from the main base and uses a manually
operated generator to periodically recharge batteries, the bulk of the
antenna stations that will comprise the full array will be located at
points farther removed from the main base and will therefore require
fully autonomous energy sources.  For the autonomous station
pathfinder, we developed a solar charging system to power the station,
and we are investigating small wind turbines for future
\albatros\ stations.

@@

We have been unable to retrieve data from the single autonomous
station pathfinder on Marion (\S\ref{s:autonomous}) because of
COVID-19 restrictions and limited bandwidth to the island, although we
have confirmed that the solar power system has successfully met the
power demands of the antenna station.  To demonstrate baseband writing
capability in the absence of Marion data, we instead present a short
observation that was taken by a separate, single-element
\albatros\ pathfinder station that operated from the McGill Arctic
Research Station (MARS; \ang{79;26;}N, \ang{90;46;}W) on Axel Heiberg
Island, Nunavut, during July~2019.  The RF signal chain of the MARS
installation had the same architecture as that shown in
\autoref{Fig:albatros1_schem}, but the power was supplied by manually
charged batteries, rather than a solar power system.  The MARS RFI
environment is somewhat noisier than that of Marion Island, and
because the data were taken during the height of Arctic summer, the
ionospheric plasma cutoff frequency is higher than the typical cutoff
reflected in the Marion data.  Despite the differences in the power
configuration and RFI environment, the MARS installation is suitable
for illustrating baseband data writing.
